Simulation study in rectangular domain

### 2024-04-19 R/simulation_study/simulation_study_fjords.Rmd
parallelization in simulation_study_rect + other stuff
changes in simulation study

# Import libraries and data

@@

# Get land polygons

@@

```{r fix time steps}
#observation times
Tmax=3*24 #max time in hour

@@

#constant tau
ftau_constant=function(Dshore,theta,tau=1) {
  return (tau)
}

@@

nu=4

@@

```{r}
sp_df=5
#number of points
n <- 500

@@

## Simulation with measurement errors and high frequency

@@

```{r high frequency samples no error (hf ne)}

#initialize data frame
data_fjords_hf_ne=data.frame("y1"=NA,"y2"=NA,"time"=NA,"p1"=NA,"p2"=NA,"ID"=NA)

for (i in 1:n_samples) {

  res=sim_theta_CRCVM(ftau_constant,fomega,nu,log_sigma_obs=log(sigma_obs_0),v0,x0[i,],times_hf,land=land,verbose=FALSE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_shore=res$shore[,c("p1","p2")]
  data_fjords_hf_ne=rbind(data_fjords_hf_ne,cbind(data_sim,data_shore))
}

data_fjords_hf_ne=data_fjords_hf_ne[-1,]

@@

```{r percentage of hf ne samples that reached land}
count=0
for (id in unique(data_fjords_hf_ne$ID)){
  sub_data=data_fjords_hf_ne[data_fjords_hf_ne$ID==id,]
  if (nrow(sub_data) < n_hf) {
    print(id)
    count=count+1
  }
}
cat(count/n_samples*100,"percent of the samples reached land")
```

### Plot samples

```{r plot hf ne samples}

plot_samples_fjords_hf_ne=ggplot()+geom_sf(data=land$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_fjords_hf_ne[data_fjords_hf_ne$ID==2,],mapping=aes(y1,y2),size=0.1)+
  geom_point(data=data_fjords_hf_ne[data_fjords_hf_ne$ID==2,],mapping=aes(p1,p2),size=0.01,col="red",shape=5)+
  geom_path(data=data_fjords_hf_ne[data_fjords_hf_ne$ID==2,],mapping=aes(y1,y2),linewidth=0.1)+
  geom_point(data = data_fjords_hf_ne[data_fjords_hf_ne$ID==2,] %>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4,col = "blue")+
  xlab("x") + ylab("y")

## sImulate with high frequency and measurement error

```{r generate high frequency samples with error (hf e)}

#initialize data frame
data_fjords_hf_e=data.frame("y1"=NA,"y2"=NA,"time"=NA,"p1"=NA,"p2"=NA,"ID"=NA)

for (i in 1:n_samples) {

  res=sim_theta_CRCVM(ftau_constant,fomega,nu,log_sigma_obs=log(sigma_obs_1),v0,x0[i,],times_hf,land=land,verbose=FALSE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_shore=res$shore[,c("p1","p2")]
  data_fjords_hf_e=rbind(data_fjords_hf_e,cbind(data_sim,data_shore))
}

data_fjords_hf_e=data_fjords_hf_e[-1,]

### Percentage of samples that reached land
```{r percentage of hf e samples that reached land}
count=0
for (id in unique(data_fjords_hf_e$ID)){
  sub_data=data_fjords_hf_e[data_fjords_hf_e$ID==id,]
  if (nrow(sub_data) < n_hf) {
    print(id)
    count=count+1
  }
}
cat(count/n_samples*100,"percent of the samples reached land")
```

### Plot one sample

```{r plot hf e samples}

plot_samples_fjords_hf_e=ggplot()+geom_sf(data=land$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_fjords_hf_e[data_fjords_hf_e$ID==2,],mapping=aes(y1,y2),size=0.1)+
   geom_path(data=data_fjords_hf_e[data_fjords_hf_e$ID==2,],mapping=aes(y1,y2),linewidth=0.1)+
  geom_point(data = data_fjords_hf_e[data_fjords_hf_e$ID==2,] %>% filter(!duplicated(ID)),
             aes(x = y1, y = y2 ), shape = 3, size = 4,col = "red")+
    xlab("x") + ylab("y")

@@

```{r compute covariates}

@@

add_covs=function(data) {
  #data has at least columns "time","y1" "y2", "p1","p2"
  #this function returns a dataframe with three mùore columns : "theta", "DistanceShore" and "ExpShore"

@@

data_fjords_hf_ne=add_covs(data_fjords_hf_ne)

data_fjords_hf_e=add_covs(data_fjords_hf_e)
```

@@

```{r estimate from hf e samples with cov ExpShore}

#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,ExpShore,k=5,bs="cs"))
par0 <- c(0,0,1,4,0)

crcvm_fjords_hf_ne1 <- SDE$new(formulas = formulas,data = data_fjords_hf_ne[data_fjords_hf_ne$ID==2,],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2","nu","tau"),
                      other_data=list("H"=H0_hf))
crcvm_fjords_hf_ne1$fit(method="BFGS")
```

#### Plot estimated smooth parameter omega in 3D

@@

@@

@@

atw=list("ExpShore"=fExpShore,"AngleNormal"=fangle)
sim_data=crcvm_fjords_hf_ne1$simulate(z0=x0[1,],data=crcvm_fjords_hf_ne1$data(),atw=atw,land=land)

>>>>

# SET UP
## Import libraries and data

@@

## Global variables

```{r}
#number of batch
N_BATCH=2

#number of individual tracks per batch
N_ID=3

#duration of each track in hour
TMAX=0.5*24

#degree of freedom in tensor splines
SP_DF=5
```

## Get land polygons

@@

```{r fix time steps}

@@

#constant tau
ftau_constant=function(cov_data,tau=1,sigma_re=0.1) {
  b=rnorm(1,mean=0,sd=sigma_re)
  return (tau+b)
}

@@

#constant nu
fnu_constant=function(cov_data,nu=4,sigma_re=0.3) {
    b=rnorm(1,mean=0,sd=sigma_re)
  return (nu+b)
}

@@

```{r}
sp_df=5
#number of points
n <- 1000

@@

### Define spline smooth functions for omega based on the approximation

```{r spline smooth omega definition}

fomega_splines=function(cov_data) {

  if (is.null(cov_data$ExpShore)){
    omega=predict(m1,newdata=cov_data)
  }
  else if (is.null(cov_data$DistanceShore)){
    omega=predict(m2,newdata=cov_data)
  }

  return(as.numeric(omega))
}

## Simulation with low measurement errors and high frequency

### Generate sample

```{r parallel generate hf low errors samples}

#setup parallel backend to use many processors
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)

data_fjords_hf_ne <- foreach(i=1:N_BATCH,.packages=c("progress","MASS","sf","foreach")) %dopar% {
  data=foreach (i=1:N_ID,.combine='rbind') %do% {

    res=sim_theta_CRCVM(ftau_constant,fomega,fnu_constant,log_sigma_obs=log(sigma_obs_0),v0,x0[i,],times_hf,land=land,verbose=FALSE)

    data_sim=res$sim
    data_sim$ID=factor(rep(i,length(data_sim$y1)))
    data_shore=res$shore[,c("p1","p2")]
    cbind(data_sim,data_shore)
  }
  data 
}

#stop cluster
stopCluster(cl)

names(data_fjords_hf_ne)=paste("BATCH",1:N_BATCH,sep="")
```

```{r percentage of low error samples that reached land}
count=0
for (k in 1:length(names(data_rect_hf_ne))) {

  data=data_fjords_hf_ne[[paste("BATCH",k,sep="")]]
  for (id in unique(data$ID)) {
    sub_data=data[data$ID==id,]
    if (nrow(sub_data) < n_hf) {
      count=count+1
      cat(paste("BATCH",k,sep=""),":",id,"\n",sep=" ")
    }
  }
}
cat(count/(N_BATCH*N_ID)*100,"percent of the samples reached land")
```

### plot one batch

```{r plot one batch low error samples}

plot_samples_fjords_hf_ne=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_fjords_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_path(data=data_fjords_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_point(data = data_fjords_hf_ne$BATCH1%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

## Simulation with high measurement error and high frequency

@@

```{r parallel generate hf high error samples, cache=TRUE,results='hide'}
#setup parallel backend to use many processors
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)

data_fjords_hf_e <- foreach(i=1:N_BATCH,.packages=c("progress","MASS","sf","foreach")) %dopar% {
  data=foreach (i=1:N_ID,.combine='rbind') %do% {

    res=sim_theta_CRCVM(ftau_constant,fomega,fnu_constant,log_sigma_obs=log(sigma_obs_1),v0,x0[i,],times_hf,land=land,verbose=FALSE)

    data_sim=res$sim
    data_sim$ID=factor(rep(i,length(data_sim$y1)))
    data_shore=res$shore[,c("p1","p2")]
    cbind(data_sim,data_shore)
  }
  data 
}

#stop cluster
stopCluster(cl)

#change names of the list elements
names(data_fjords_hf_e)=paste("BATCH",1:N_BATCH,sep="")
```

@@

```{r percentage of high error samples that reached land}
count=0
for (k in 1:length(names(data_fjords_hf_e))) {

  data=data_fjords_hf_e[[paste("BATCH",k,sep="")]]
  for (id in unique(data$ID)) {
    sub_data=data[data$ID==id,]
    if (nrow(sub_data) < n_hf) {
      count=count+1
      cat("BATCH",k,id,"\n",sep="")
    }
  }
}
cat(count/(N_BATCH*N_ID)*100,"percent of the samples reached land")
```

### plot one sample

```{r plot one batch of high error samples}

plot_samples_fjords_hf_e=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_fjords_hf_e$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_path(data=data_fjords_hf_e$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_point(data = data_fjords_hf_e$BATCH1%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

@@

```{r compute covs}

@@

add_covs=function(data) {
  #data has at least columns "time","y1" "y2", "p1","p2"
  #this function new data frame with three more columns "theta", "DistanceShore" and "ExpShore"

@@

for (k in 1:N_BATCH) {

  #dataframes for this batch
  df_ne=data_fjords_hf_ne[[paste("BATCH",k,sep="")]]
  df_e=data_fjords_hf_e[[paste("BATCH",k,sep="")]]

  #add covariates columns
  df_ne=add_covs(df_ne)
  df_e=add_covs(df_e)

  #change in the list of batch
  data_fjords_hf_ne[[paste("BATCH",k,sep="")]]=df_ne
  data_fjords_hf_e[[paste("BATCH",k,sep="")]]=df_e
}

@@

```{r estimate te ExpShore for low error samples,cache=TRUE}

#setup parallel backend to use many processors
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)

crcvm_fjords_hf_ne1<- foreach(i=1:N_BATCH,.packages=c("smoothSDE")) %dopar% {
  #estimation with tensor splines
  formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,ExpShore,k=SP_DF,bs="cs"))
  par0 <- c(0,0,1,1,0)
  m <- SDE$new(formulas = formulas,data = data_fjords_hf_ne[[paste("BATCH",k,sep="")]],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H0_hf))
  m$fit(method="BFGS")
  m 
}

#stop cluster
stopCluster(cl)

names(crcvm_fjords_hf_ne1)=paste("BATCH",1:N_BATCH,sep="")

#### Plot estimated smooth parameter omega in 3D for one batch

@@

#### Violin plot of the estimated spline coefficients for one batch

@@

#### Violin plot of the estimated spline coefficients for all batches

```{r violin plots spline coeff te ExpShore all batches with low measurement error}

n_coeffs=length(sp_coeff_ExpShore)
coeffs_df=data.frame("estimate"=NA,"coeff_name"=NA)

for (k in 1:N_BATCH) {
  m<-crcvm_fjords_hf_ne1[[paste("BATCH",k,sep="")]]
  coeffs=m$coeff_re()
  coeff_names=rownames(coeffs)
  coeff_values=as.numeric(coeffs)
  sub_index=grepl("omega",coeff_names)
  sub_coeffs_df=data.frame("estimate"=coeff_values[sub_index],"coeff_name"=factor(coeff_names[sub_index]))
  coeffs_df=rbind(coeffs_df,sub_coeffs_df)
}

#remove initialization row
coeffs_df=coeffs_df[-1,]

p=ggplot()+geom_violin(data=coeffs_df,aes(x=coeff_name,y=estimate,fill=coeff_name))+
                xlab(" ")+labs(fill = "Estimated coefficients")+
                theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
                geom_point(data = true_coeff, aes(x = name, y = value), color = "red", size = 1,shape=4)

p=ggplotly(p)
saveWidget(p, file ="crcvm_fjords_hf_ne1/all_batches_coeffs_crcvm_fjords_hf_ne1.png")

@@

atw=list("ExpShore"=fExpShore,"AngleNormal"=fangle)
sim_data=crcvm_rect_hf_ne1$BATCH1$simulate(z0=x0[1,],data=crcvm_rect_hf_ne1$BATCH1$data(),atw=atw,land=domain)

### 2024-04-19 R/simulation_study/simulation_study_rect.Rmd
changes in simulation study

ggsave("all_batches_coeffs_crcvm_rect_hf_ne1.png",plot=p,path="crcvm_rect_hf_ne1")

@@

ggsave("all_batches_coeffs_crcvm_rect_hf_ne2.png",plot=p,width=10,height=5,path="crcvm_rect_hf_ne2")

@@

ggsave("all_batches_coeffs_crcvm_rect_hf_e1.png",plot=p,width=10,height=5,path="crcvm_rect_hf_e1")

ggplotly(p)

@@

ggsave("all_batches_coeffs_crcvm_rect_hf_e2.png",plot=p,width=10,height=5,path="crcvm_rect_hf_e2")
```

>>>>

p=ggplotly(p)
saveWidget(p, file ="crcvm_rect_hf_ne1/all_batches_coeffs_crcvm_rect_hf_ne1.html")

@@

p=ggplotly(p)
saveWidget(p, file ="crcvm_rect_hf_ne2/all_batches_coeffs_crcvm_rect_hf_ne2.html")

@@

p=ggplotly(p)
saveWidget(p, file ="crcvm_rect_hf_e1/all_batches_coeffs_crcvm_rect_hf_e1.html")

@@

p=ggplotly(p)
saveWidget(p, file ="crcvm_rect_hf_e2/all_batches_coeffs_crcvm_rect_hf_e2.html")
```
